---
title: "Group_U report"
author: "Group_U"
date: "2025-02-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Required Packages

```{r include=FALSE, results='hide'}
packages = (c("ipumsr", "dplyr", "ggplot2", "tidyr", "here", "rmdformats", "srvyr", "DiagrammeR", "gridExtra", "survey", "Matrix", "haven", "gtsummary", "pscl","lmtest", "sandwich","zoo","stargazer","gt","modelsummary", "here"))

## Now load or install&load all
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

# Data

## Data Import and Initial Inspection

```{r setup, include=True}
#set working documents
current_dir <- dirname(rstudioapi::getSourceEditorContext()$path)

#read data
ddi <- read_ipums_ddi(file.path(current_dir, "usa_00007.xml"))
data <- read_ipums_micro(ddi, data_file = file.path(current_dir, "usa_00007.dat"))

#check first 4 rows
head(data)

count(data,SEX) %>%
  mutate(n = format(n, big.mark = ",", scientific = FALSE))

names(data)
```

## Causal Relationship Visualization (**DAGs)**

```{r setup, include=True}

dag <- grViz("
  digraph DAG {
   age -> fertility
   age -> workinghours
   education -> fertility
   education -> workinghours
   workinghours -> income -> fertility
  }
")
dag
```

## Visualizing data before cleaning

```{r setup, include=True}

# 1. Depandent variable_number of children distribution
p1 <- ggplot(data, aes(x = NCHILD)) +
  geom_bar(fill = "skyblue") +
  theme_minimal() +
  labs(title = "Distribution of Number of Children",
       x = "Number of Children",
       y = "Count")

# 2. Main indepandent variable_working hours distribution
p2 <- ggplot(data, aes(x = UHRSWORK)) +
  geom_histogram(binwidth = 5, fill = "lightgreen") +
  theme_minimal() +
  labs(title = "Distribution of Working Hours",
       x = "Working Hours per Week",
       y = "Count")

# 3. Income distribution
p3 <- ggplot(data, aes(x = INCTOT)) +
  geom_histogram(fill = "pink") +
  scale_x_log10() +  # consider income is distributed as right skewed
  theme_minimal() +
  labs(title = "Distribution of Income (log scale)",
       x = "Income",
       y = "Count")

# 4. Age distribution
p4 <- ggplot(data, aes(x = AGE)) +
  geom_histogram(binwidth = 1, fill = "lightyellow") +
  theme_minimal() +
  labs(title = "Distribution of Age",
       x = "Age",
       y = "Count")

# Combined graphs
grid.arrange(p1, p2, p3, p4, ncol = 2)

```

## Data cleaning

```{r setup, include=True}

# check unprocessed data
cat("\n=== summary unprocessed data ===\n")
print("income summary：")
summary(data$INCTOT)
print("\n working hours summary：")
summary(data$UHRSWORK)
print("\n gender distribution：")
table(data$SEX)
print("\n age distribution：")
summary(data$AGE)
print("education summary：")
summary(data$EDUC)

# calculate IQR bounds for "UHRSWORK"
Q1_work <- quantile(data$UHRSWORK, 0.25, na.rm = TRUE)
Q3_work <- quantile(data$UHRSWORK, 0.75, na.rm = TRUE)
IQR_work <- Q3_work - Q1_work
lower_bound_work <- Q1_work - 1.5 * IQR_work
upper_bound_work <- Q3_work + 1.5 * IQR_work

# calculate IQR bounds for "INCTOT"
Q1_inc <- quantile(data$INCTOT, 0.25, na.rm = TRUE)
Q3_inc <- quantile(data$INCTOT, 0.75, na.rm = TRUE)
IQR_inc <- Q3_inc - Q1_inc
upper_bound_inc <- Q3_inc + 1.5 * IQR_inc

# filter data using IQR method and keep only females
data_cleaned <- data %>%
  filter(!is.na(INCTOT)) %>%  #remove NA values
 
  filter(INCTOT >= 0 & INCTOT < upper_bound_inc) %>%
  #↑There are negative incomes existing if we use "greater than lower bounds" here, therefore we use "equal and greater than 0" instead
  
  filter(AGE >= 18 & AGE <= 49) %>%
  filter(UHRSWORK > 0 & UHRSWORK <= upper_bound_work) %>%
  filter(SEX == 2)

# create log(income) and transfer EDUC to dummy variable
data_cleaned <- data_cleaned %>%
  mutate(
    log_income = ifelse(INCTOT <= 0, NA, log(INCTOT)),
    EDUC_Medium = ifelse(EDUC > 2 & EDUC <= 6, 1, 0),
    EDUC_High = ifelse(EDUC > 6, 1, 0)
  )
  

# check processed data
cat("\n=== summary processed data ===\n")
print("\n cleaned income summary：")
summary(data_cleaned$INCTOT)
print("\n cleaned working hours summary：")
summary(data_cleaned$UHRSWORK)
print("\n cleaned gender distribution：")
table(data_cleaned$SEX)
print("\n cleaned age distribution：")
summary(data_cleaned$AGE)
print("education summary：")
summary(data$EDUC)

# print boundaries
cat("\n=== unusual bounds ===\n")
cat("\nWork Hours bounds：")
print(paste("Lower bounds:", round(lower_bound_work, 2)))
print(paste("Upper bounds:", round(upper_bound_work, 2)))

cat("\n Income bounds：")
print(paste("Lower bound: 0"))
print(paste("Upper bound:", round(upper_bound_inc, 2)))

# check removed records
cat("\n=== removed data summary ===\n")
print(paste("Total removed records:", nrow(data) - nrow(data_cleaned)))
print(paste("Total percentage removed:", round((nrow(data) - nrow(data_cleaned))/nrow(data)*100, 2), "%"))

```

## Data quality and missing value check

```{r setup, include=True}
count(data_cleaned, SEX) %>%
  mutate(n = format(n, big.mark = ",", scientific = FALSE))

na_count <- colSums(is.na(data))
print(na_count)  # show NA values of each variable

print("cleaned data")
na_count <- colSums(is.na(data_cleaned))
print(na_count)  # show NA values of each variable
```

## Visualizing data after cleaning

```{r setup, include=True}

# 1. Depandent variable_number of children distribution_after cleaning
p1 <- ggplot(data_cleaned, aes(x = NCHILD)) +
  geom_bar(fill = "skyblue") +
  theme_minimal() +
  labs(title = "Distribution of Number of Children_after cleaning",
       x = "Number of Children",
       y = "Count") +
  theme(plot.title = element_text(size = 10))

# 2. Main indepandent variable_working hours distribution
p2 <- ggplot(data_cleaned, aes(x = UHRSWORK)) +
  geom_histogram(binwidth = 5, fill = "lightgreen") +
  theme_minimal() +
  labs(title = "Distribution of Working Hours_after cleaning",
       x = "Working Hours per Week",
       y = "Count")+
  theme(plot.title = element_text(size = 10))

# 3. Income distribution
p3 <- ggplot(data_cleaned, aes(x = INCTOT)) +
  geom_histogram(fill = "pink") +
  scale_x_log10() +  # consider income is distributed as right skewed
  theme_minimal() +
  labs(title = "Distribution of Income (log scale)_after cleaning",
       x = "Income",
       y = "Count")+
  theme(plot.title = element_text(size = 10))

# 4. Age distribution
p4 <- ggplot(data_cleaned, aes(x = AGE)) +
  geom_histogram(binwidth = 1, fill = "lightyellow") +
  theme_minimal() +
  labs(title = "Distribution of Age_after cleaning",
       x = "Age",
       y = "Count")+
  theme(plot.title = element_text(size = 10))

# Combined graphs
grid.arrange(p1, p2, p3, p4, ncol = 2)

```

## Sampling design and weight variable processing

```{r}

data_cleaned <- data_cleaned %>%
  mutate(weight = as.numeric(PERWT)) 

design <- svydesign(ids = ~1, weights = ~weight, data = data_cleaned)
```

## Data conversion and descriptive statistics table

```{r setup, include=True}

data_cleaned <- data_cleaned %>%
  mutate(
    UHRSWORK = as.numeric(unlist(UHRSWORK)),
    INCTOT = as.numeric(unlist(INCTOT)),
    AGE = as.numeric(unlist(AGE)),
    # create age_groups
    age_group = case_when(
      AGE < 30 ~ "young",
      AGE < 40 ~ "middle",
      TRUE ~ "older"
    ),
    age_group = factor(age_group, levels = c("young", "middle", "older")),
    # create education level groups
    EDUC = as.numeric(unlist(EDUC)),
    EDUC_Medium = ifelse(EDUC > 2 & EDUC <= 6, 1, 0),
    EDUC_High = ifelse(EDUC > 6, 1, 0),
    # factor education level
    education = case_when(
      EDUC <= 2 ~ "Low",
      EDUC > 2 & EDUC <= 6 ~ "Medium",
      EDUC > 6 ~ "High"
    ),
    education = factor(education, levels = c("Low", "Medium", "High"))
  )

# output descriptive table
summary_table <- data_cleaned %>%
  tbl_summary(
    include = c(UHRSWORK, INCTOT, AGE, education),
    by = age_group,
    label = list(
      UHRSWORK ~ "Weekly working hours",
      INCTOT ~ "Income",
      AGE ~ "Age",
      education ~ "Education Level"
    ),
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    digits = list(all_continuous() ~ 2)
  ) %>%
  add_p() %>%
  add_overall() %>%
  modify_header(label ~ "**Characteristic**") %>%
  modify_footnote(
    all_stat_cols() ~ "Mean (SD); n (%)"
  )

# show table
summary_table
```

## Relationship between fertility and age (non-linear)

```{r setup, include=True}
# 1. sampling data
set.seed(123)
sample_size <- min(10000, nrow(data_cleaned))
sampled_data <- data[sample(nrow(data_cleaned), sample_size), ]

# 2. plot by using sampling data set
p7 <- ggplot(sampled_data, aes(x = AGE, y = NCHILD)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "loess", color = "red") +
  theme_minimal() +
  labs(title = "Age vs Number of Children",
       x = "Age",
       y = "Number of Children")

p7
```

# Model Regression

## model_a (no income, no age-square)

```{r setup, include=True}

# create poisson model
model_a <- glm(NCHILD ~ UHRSWORK + AGE + 
            EDUC_Medium + EDUC_High, 
            family = poisson(link = "log"), 
            data = data_cleaned)

robust_se_a <- coeftest(model_a, vcov = vcovHC(model_a, type = "HC0")) #robust standard error

# output model result
stargazer(model_a, 
          type = "text",
          title = "Results -  Model A_Basic Model ",
          digits = 3)

summary(model_a)
print(robust_se_a)
pR2(model_a)
pR2_a <- pR2(model_a)["McFadden"]
print(pR2_a)

# Residual Plots
residuals <- data.frame(
  Fitted = fitted(model_a),
  Residuals = residuals(model_a)
)


ggplot(residuals, aes(x = Fitted, y = Residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Residual Plot:  Basic model A",
    x = "Fitted Values",
    y = "Residuals"
  )


# Scatter Plot of Working Hours vs Number of Children
ggplot(data_cleaned, aes(x = UHRSWORK, y = NCHILD)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", color = "red") +
  labs(x = "Weekly Work Hours", y = "Number of Children", 
       title = "Relationship between Work Hours and Number of Children")

```

## model_b (add age-square)

```{r setup, include=True}

# scatter plot
ggplot(data_cleaned, aes(x = UHRSWORK, y = NCHILD)) +
  geom_point(alpha = 0.1) + 
  geom_smooth(method = "lm", color = "red") +  
  labs(title = "Relationship between Working Hours and Number of Children",
       x = "Weekly Working Hours",
       y = "Number of Children") +
  theme_minimal()

# create model with "AGE^2"
model_b <- glm(NCHILD ~ UHRSWORK + AGE + I(AGE^2) + EDUC_Medium + EDUC_High, 
               data = data_cleaned, 
               family = poisson(link = "log"))

robust_se_b <- coeftest(model_b, vcov = vcovHC(model_b, type = "HC0")) #robust standard error

overdispersion_rate <- sum(residuals(model_b, type = "pearson")^2) / model_b$df.residual #overdispersion rate by calculating the ratio of the sum of squared Pearson residuals to the residual degrees of freedom

# output model result
stargazer(model_b, 
          type = "text",
          title = "Results -  Model B_Adding Nonlinearity in Age ",
          digits = 3)

# check distribution of education level
print("Education Level Distribution:")
print(table(data_cleaned$edu_level))

# Residual Plots
residuals <- data.frame(
  Fitted = fitted(model_b),
  Residuals = residuals(model_b)
)

ggplot(residuals, aes(x = Fitted, y = Residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Residual Plot:  Model B_Adding Nonlinearity in Age ",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()

# output summary
print("Model Summary:")
summary(model_b)
print(robust_se_b)
pR2(model_b)
pR2_b <- pR2(model_b)["McFadden"]
print(pR2_b)
print(overdispersion_rate)
```

## model_c (add "income")

```{r setup, include=True}

# scatter plot
ggplot(data_cleaned, aes(x = UHRSWORK, y = NCHILD)) +
  geom_point(alpha = 0.1) +  
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Relationship between Working Hours and Number of Children",
       x = "Weekly Working Hours",
       y = "Number of Children") +
  theme_minimal()



# create the model with "log(income) based on model_b
model_c <- glm(NCHILD ~ UHRSWORK + AGE + I(AGE^2) +  log_income + EDUC_Medium + EDUC_High, 
               data = data_cleaned, 
               family = poisson(link = "log"))
#robust se
robust_se_c <- coeftest(model_c, vcov = vcovHC(model_c, type = "HC0")) #robust standard error


# output result
stargazer(model_c, 
          type = "text",
          title = "Regression Results -  Model C_Testing for Mediation by Income",
          digits = 3)

# check distribution of education level
print("Education Level Distribution:")
print(table(data_cleaned$edu_level))

# residual plot
residuals <- data.frame(
  Fitted = fitted(model_c),
  Residuals = residuals(model_c)
)

ggplot(residuals, aes(x = Fitted, y = Residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Residual Plot: Model C_Testing for Mediation by Income",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()

# summary model
print("Model Summary:")
summary(model_c)
print(robust_se_c)
pR2(model_c)
pR2_c <- pR2(model_c)["McFadden"]
print(pR2_c)

```

## Comparison for Three Models

```{r setup, include=TRUE}

# list models
models <- list(
  "Model_basic" = model_a,
  "Model_age-square" = model_b,
  "Model_with_income" = model_c
)

# create a new row
new_row <- tibble(
  term = "R²",
  `Model_basic` = sprintf("%.3f", pR2(model_a)["McFadden"]),
  `Model_age-square` = sprintf("%.3f", pR2(model_b)["McFadden"]),
  `Model_with_income` = sprintf("%.3f", pR2(model_c)["McFadden"])
)

modelsummary(models, 
             estimate = "estimate",
             statistic = "({std.error})", 
             stars = TRUE,
             coef_rename = c(
               "UHRSWORK" = "Working Hours",
               "AGE" = "Age",
               "I(AGE^2)" = "Age Squared",
               "edu_levelMedium" = "Education: Medium",
               "edu_levelHigh" = "Education: High",
               "INCWAGE" = "Income"
             ),
             add_rows = new_row,
             title = 'Impact of Working Hours on Fertility: Model Comparison')

```

## **Heterogeneity Test**

```{r setup, include=True}

#interaction model
model_interaction <- glm(NCHILD ~ UHRSWORK * EDUC_Medium + UHRSWORK * EDUC_High + AGE + I(AGE^2), 
                         data = data_cleaned, family = poisson(link = "log"))

robust_se_interaction <- coeftest(model_interaction, vcov = vcovHC(model_interaction, type = "HC0"))
summary(model_interaction)
pR2(model_interaction)

# residual plot
residuals_interaction <- data.frame(
  Fitted_interaction = fitted(model_interaction),
  Residuals_interaction = residuals(model_interaction)
)
ggplot(residuals_interaction, aes(x =  Fitted_interaction, y = Residuals_interaction)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Residual Plot: model_interaction",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()



```

